---
title: "Final Report"
author: "Lindsay Knupp"
date: "2024-05-14"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dataset



## Simulation Study 

We were interested in understanding how our data affected the optimal choice of $k$ in the $k$ -nearest neighbors algorithm. We already experienced some variability when running our model through on different computers. Therefore, we wanted to see if more simulated datasets would produce the same variablility. 

To replicate our $13$ predictors and $2$ response variables, we used a package called `faux` to simulate our numerical predictors from a multivariate normal distribution. 

We used a standard normal distribution to replicate our $13$ predictors. 

```{r,warning=FALSE,message=FALSE,echo=FALSE}
library(ggplot2)
library(readr)

results = read_csv("~/hospital-costs/lindsayknupp/sim_study_results.csv")

ggplot(results) +
  geom_histogram(aes(x=k),bins=15) +
  ggtitle("Optimal k over 200 datasets") +
  xlab("k") + 
  ylab("count")

ggplot(results) +
  geom_point(aes(x=k,y=error)) + 
  ggtitle("Error rates vs optimal k")

ggplot(results) + 
  geom_histogram(aes(x=var),bins=15)

ggplot(results) +
  geom_point(aes(x=k,y=var)) +
  ggtitle("Error variance vs optimal k")

```

